---
title: "Forest structure variables"
author: "Gene Estrada"
date: "2024-06-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(PerformanceAnalytics)
library(cowplot)

#select stand metrics
stand_mets <- read.csv(file = "data/select.stand.mets.csv", header = TRUE)

#ordering habitat by elevation
stand_mets$habitat <- factor(stand_mets$habitat, 
                                       levels = c("Peat Swamp", "Freshwater Swamp", "Alluvial Bench", 
                                                "Lowland Sandstone", "Lowland Granite", 
                                                "Upland Granite", "Montane"),
                                       ordered = TRUE)

stand_mets$partition <- factor(stand_mets$partition, 
                                       levels = c("PS1", "FS1", "AB1", "AB2",
                                                "LS1", "LS2", "LG1", "LG2",
                                                "UG1", "UG2", "MO1", "MO2"),
                                       ordered = TRUE)
#all tree metrics
tree_mets <- read.csv(file = "data/all.tree.mets.csv", header = TRUE)

#ordering habitat by elevation
tree_mets$habitat <- factor(tree_mets$habitat, 
                                       levels = c("Peat Swamp", "Freshwater Swamp", "Alluvial Bench", 
                                                "Lowland Sandstone", "Lowland Granite", 
                                                "Upland Granite", "Montane"),
                                       ordered = TRUE)

tree_mets$partition <- factor(tree_mets$partition, 
                                       levels = c("PS1", "FS1", "AB1", "AB2",
                                                "LS1", "LS2", "LG1", "LG2",
                                                "UG1", "UG2", "MO1", "MO2"),
                                       ordered = TRUE)
#all stand metrics
all_stand_mets <- read.csv(file = "data/all.stand.mets.csv", header = TRUE)

#changing default setting in ggplot to center plot titles
theme_update(plot.title = element_text(hjust = 0.5))

ft_colors <- c("Peat Swamp" = "#401D00", 
                      "Freshwater Swamp" = "#643E00", 
                      "Alluvial Bench" = "#896007", 
                      "Lowland Sandstone" = "#AF8228", 
                      "Lowland Granite" = "#D3A43B", 
                      "Upland Granite" = "#F4C444",
                      "Montane" = "#FFDD2D")
```

## Correlations between variables
```{r warning=FALSE}
chart.Correlation(stand_mets[,c(5:16,21:29)], histogram = TRUE, pch = 19)
```

## On- and Off-trail scans
by forest type
```{r}
table(stand_mets$habitat, stand_mets$on.off.trail)
```

## Plotting forest structure metrics by forest type

### Maximum tree height
```{r}
#global settings
title_size = 10
axis_title_size = 9
axis_label_size = 9

#distribution
hist(stand_mets$max.height, breaks = 25, las = 1,
     main = "Max. Tree Heights", xlab = "tree h (m)")

#summary stats
summary(stand_mets$max.height)

#max height by forest type
plot1 <- ggplot(stand_mets, aes(habitat, max.height, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree h max.", x = "", y = "height (m)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size),
        axis.text.y = element_text(size = axis_label_size
                                   #,color = ft_colors
                                   )) + 
  coord_flip() +
  scale_color_manual(values = ft_colors)

#by partition 
ggplot(stand_mets, aes(partition, max.height)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree h max", x = "", y = "height (m)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip()

#by elevation
p1 <- ggplot(stand_mets, aes(altitude, max.height, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) + 
  labs(title = "tree h max", x = "", y = "height (m)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Mean tree height
```{r}
#distribution
hist(stand_mets$mean.tree.h, breaks = 25, las = 1,
     main = "Average Tree Heights", xlab = "mean tree h (m)")

#summary stats
summary(stand_mets$mean.tree.h)

#by forest type
plot2 <- ggplot(stand_mets, aes(habitat, mean.tree.h, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree h mean", x = "", y = "height (m)") + 
  coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, mean.tree.h)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree h mean", x = "", y = "mean tree h (m)") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p2 <- ggplot(stand_mets, aes(altitude, mean.tree.h, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "tree h mean", x = "", y = "height (m)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Canopy relief ratio
```{r}
#distribution 
hist(stand_mets$CRR.rho, breaks = 25, las = 1,
     main = "Canopy Relief Ratio", xlab = "CRR")

#summary stats
summary(stand_mets$CRR.rho)

#by forest type
plot3 <- ggplot(stand_mets, aes(habitat, CRR.rho, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "canopy relief ratio", x = "", y = "CRR") + 
  coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition - sizable differences between all partitions except UG
ggplot(stand_mets, aes(partition, CRR.rho)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "canopy relief ratio", x = "", y = "CRR") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p3 <- ggplot(stand_mets, aes(altitude, CRR.rho, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "canopy relief ratio", x = "", y = "CRR") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Standard deviation of tree heights
```{r}
#distribution 
hist(stand_mets$sd.r, breaks = 25, las = 1,
     main = "Std. Dev. Tree Heights", xlab = "sd tree h (m)")

#summary stats
summary(stand_mets$sd.r)

#by forest type
plot4 <- ggplot(stand_mets, aes(habitat, sd.r, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree h sd", x = "", y = "height sd (m)") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size),
        axis.text.y = element_text(size = axis_label_size)) + 
  scale_color_manual(values = ft_colors)

#by partition - big difference between AB1 and AB2
ggplot(stand_mets, aes(partition, sd.r)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree h sd", x = "", y = "sd tree h (m)") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p4 <- ggplot(stand_mets, aes(altitude, sd.r, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "tree h sd", x = "", y = "height sd (m)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Mean DBH
```{r}
#distribution
hist(stand_mets$mean.dbh, breaks = 25,
     main = "Avg. Tree DBH", xlab = "dbh (cm)")

#summary stats
summary(stand_mets$mean.dbh)

#by forest type
plot5 <- ggplot(stand_mets, aes(habitat, mean.dbh, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree dbh", x = "", y = "diameter (cm)") + 
  coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, mean.dbh)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree dbh", x = "", y = "diameter (cm)") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size))

#by elevation
p5 <- ggplot(stand_mets, aes(altitude, mean.dbh, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "tree dbh", x = "", y = "diameter (cm)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Basal area
```{r}
#distribution - some outliers on the upper end, pulling the mean higher
hist(stand_mets$basal.area, breaks = 25,
     main = "Tree Basal Area", xlab = expression(m^2/ha))

#summary stats
summary(stand_mets$basal.area)

#by forest type
plot6 <- ggplot(stand_mets, aes(habitat, basal.area, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "basal area", x = "", y = "sq. m /ha") + 
  coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, basal.area)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "basal area", x = "", y = expression(m^2/ha)) + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p6 <- ggplot(stand_mets, aes(altitude, basal.area, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "basal area", x = "", y = expression(m^2/ha)) + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Tree density 
```{r}
#distribution
hist(stand_mets$stand.dens, breaks = 25,
     main = "Tree Density", xlab = "trees/ha")

#summary stats
summary(stand_mets$stand.dens)

#by forest type
plot7 <- ggplot(stand_mets, aes(habitat, stand.dens, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree density", x = "", y = "n trees/ha") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size),
        axis.text.y = element_text(size = axis_label_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, stand.dens)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree density", x = "", y = "trees/ha") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p7 <- ggplot(stand_mets, aes(altitude, stand.dens, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "tree density", x = "", y = "n trees/ha") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Stem volume 
```{r}
#distribution - outliers on upper end
hist(stand_mets$stem.vol, breaks = 25,
     main = "Tree Volume", xlab = expression(m^3/ha))

#summary stats - mean quite a bit higher than median
summary(stand_mets$stem.vol)

#by forest type
plot8 <- ggplot(stand_mets, aes(habitat, stem.vol, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree volume", x = "", y = "cubic m/ha") + 
  coord_flip() + 
  theme_classic() +
  theme(axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, stem.vol)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "tree volume", x = "", y = expression(m^3/ha)) + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p8 <- ggplot(stand_mets, aes(altitude, stem.vol, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "tree volume", x = "", y = expression(m^3/ha)) + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Number of points below 2 m
```{r}
#distribution - seems to be one outlier (BC 2.19)
hist(stand_mets$pts.below.2m, breaks = 25,
     main = "Terrestrial Vegetation", xlab = "n pts. < 2 m")

#summary stats
summary(stand_mets$pts.below.2m)

#by forest type
plot9 <- ggplot(stand_mets, aes(habitat, pts.below.2m, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "terrestrial veg.", x = "", y = "n mil. pts. < 2 m") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.y = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors) +
  scale_y_continuous(labels = function(x) x / 1000000)

#by partition
ggplot(stand_mets, aes(partition, pts.below.2m)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "terr. vegetation", x = "", y = "n pts. < 2 m") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p9 <- ggplot(stand_mets, aes(altitude, pts.below.2m, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "terrestrial veg.", x = "", y = "n mil. pts. < 2 m") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none") +
  scale_y_continuous(labels = function(x) x / 1000000)
```

### zentropy
Standardized Shannon diversity metric describing vertical complexity
```{r}
#distribution
hist(stand_mets$zentropy, breaks = 25,
     main = "Vertical Complexity", xlab = "Shannon Diversity of point heights")

#summary stats
summary(stand_mets$zentropy)

#by forest type
plot10 <- ggplot(stand_mets, aes(habitat, zentropy, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "vertical pt. dist.", x = "", y = "entropy (z)") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size),
        axis.text.y = element_text(size = axis_label_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, zentropy)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "vert. pt. dist.", x = "", y = "entropy (z)") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p10 <- ggplot(stand_mets, aes(altitude, zentropy, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "vertical pt. dist.", x = "", y = "entropy (z)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Leaf Area Density
maximum value of 1m vertical bin LAD
```{r}
#distribution
hist(stand_mets$lad.max, breaks = 25,
     main = "Leaf Area Density", xlab = "LAD")

#summary stats
summary(stand_mets$lad.max)

#by forest type
plot11 <- ggplot(stand_mets, aes(habitat, lad.max, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "leaf area density", x = "", y = "LAD") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.y = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, lad.max)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "leaf area density", x = "", y = "LAD") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p11 <- ggplot(stand_mets, aes(altitude, lad.max, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "leaf area density", x = "", y = "LAD") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Rumple
Rumple index describes canopy rugosity
```{r}
#distribution
hist(stand_mets$rumple, breaks = 25, las = 1,
     main = "Canopy Rugosity", xlab = "Rumple Index")

#summary stats
summary(stand_mets$rumple)

#by forest type
plot12 <- ggplot(stand_mets, aes(habitat, rumple, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "canopy roughness", x = "", y = "rumple index") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.y = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, rumple)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "canopy roughness", x = "", y = "rumple index") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p12 <- ggplot(stand_mets, aes(altitude, rumple, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "canopy roughness", x = "", y = "rumple index") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.x = element_blank(),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Volumetric vegetation
vFRcanopy metric is the ratio of 'filled' to 'all' voxels (1 cubic m)
```{r}
#distribution
hist(stand_mets$vFRcanopy, breaks = 25, las = 1,
     main = "Volumetric Vegetation", xlab = "vegetated:all voxels")

#summary stats
summary(stand_mets$vFRcanopy)

#by forest type
plot13 <- ggplot(stand_mets, aes(habitat, vFRcanopy, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "veg. vol.", x = "", y = "vegetated:all voxels") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size),
        axis.text.y = element_text(size = axis_label_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, vFRcanopy)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "veg. vol.", x = "", y = "vegetated:all voxels") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p13 <- ggplot(stand_mets, aes(altitude, vFRcanopy, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "veg. vol.", x = "", y = "vegetated:all voxels") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Vertical rumple index
```{r}
#distribution
hist(stand_mets$vzrumple, breaks = 25, las = 1,
     main = "Vertical Rumple", xlab = "Rumple Index")

#summary stats
summary(stand_mets$vzrumple)

#by forest type
plot14 <- ggplot(stand_mets, aes(habitat, vzrumple, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "vertical pt. complex.", x = "", y = "vert. rumple index") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.y = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, vzrumple)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "vert. pt. complex.", x = "", y = "vert. rumple index") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p14 <- ggplot(stand_mets, aes(altitude, vzrumple, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "vertical pt. complex.", x = "elevation (m)", y = "vert. rumple index") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```

### Closed Gap Space
```{r}
#distribution
hist(stand_mets$ClosedGapSpace, breaks = 25, las = 1,
     main = "Closed Gap Space", xlab = "gap volume")

#summary stats
summary(stand_mets$ClosedGapSpace)

#by forest type
plot15 <- ggplot(stand_mets, aes(habitat, ClosedGapSpace, color = habitat)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "gap volume", x = "", y = "cubic m") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.text.y = element_blank(),
        legend.position = "none",
        axis.title.x = element_text(size = axis_title_size)) + 
  scale_color_manual(values = ft_colors)

#by partition
ggplot(stand_mets, aes(partition, ClosedGapSpace)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1) + 
  labs(title = "gap volume", x = "", y = "gap volume") + 
  coord_flip() + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#by elevation
p15 <- ggplot(stand_mets, aes(altitude, ClosedGapSpace, color = habitat)) + 
  geom_point() +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) +
  labs(title = "gap volume", x = "", y = expression(m^3)) + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = title_size),
        axis.title.y = element_text(size = axis_title_size),
        legend.position = "none")
```


### combining plots into a single grid

forest type plot grid
```{r}
cowplot::plot_grid(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, 
          plot10, plot11, plot12, plot13, plot14, plot15,
          nrow = 5, rel_widths = c(.42,.3,.3))

#ggsave("FS_by_FT_multiplot.jpg", plot = last_plot(), width = 7.5, height = 9.5, dpi = 500, device = "jpeg")
```

elevation plot grid
```{r}
cowplot::plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13, p14, p15, nrow = 5)

#ggsave("FS_by_elev_multiplot.jpg", plot = last_plot(), width = 7.5, height = 9.5, dpi = 500, device = "jpeg")
```


## All trees metrics

Using metrics from individual trees from all scans. Total of 2,010 trees detected by the FORTLS functions.
```{r}
#tree h
mean_h <- mean(tree_mets$h, na.rm = TRUE)
sd_h <- sd(tree_mets$h, na.rm = TRUE)

t1 <- ggplot(tree_mets, aes(h)) +
  geom_histogram(bins = 100) +
  geom_vline(aes(xintercept = mean_h), color = "red2", size = 1) +
  geom_vline(aes(xintercept = mean_h + sd_h), color = "darkred", linetype = "dashed", size = 0.8) +
  geom_vline(aes(xintercept = mean_h - sd_h), color = "darkred", linetype = "dashed", size = 0.8) +
  #annotate("text", x = 45, y = 35, label = paste("mean:", round(mean_h, 2), 
  #                                              "\nSD:", round(sd_h, 2)),
  #         hjust = 0.8, size = 4, color = "grey10") +
  labs(title = "tree height", x = "height (m)", y = "count") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#tree dbh
mean_dbh <- mean(tree_mets$dbh, na.rm = TRUE)
sd_dbh <- sd(tree_mets$dbh, na.rm = TRUE)

t2 <- ggplot(tree_mets, aes(dbh)) +
  geom_histogram(bins = 100) +
  geom_vline(aes(xintercept = mean_dbh), color = "red2", size = 1) +
  geom_vline(aes(xintercept = mean_dbh + sd_dbh), color = "darkred", linetype = "dashed", size = 0.8) +
  geom_vline(aes(xintercept = mean_dbh - sd_dbh), color = "darkred", linetype = "dashed", size = 0.8) +
  #annotate("text", x = 100, y = 92, label = paste("mean:", round(mean_dbh, 2), 
  #                                                  "\nSD:", round(sd_dbh, 2)),
  #         hjust = 0.8, size = 4, color = "grey10") +
  labs(title = "tree dbh", x = "diameter (cm)", y = "") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

#tree vol
mean_v <- mean(tree_mets$v, na.rm = TRUE)
sd_v <- sd(tree_mets$v, na.rm = TRUE)

t3 <- ggplot(tree_mets, aes(v)) +
  geom_histogram(bins = 100) +
  geom_vline(aes(xintercept = mean_v), color = "red2", size = 1) +
  geom_vline(aes(xintercept = mean_v + sd_v), color = "darkred", linetype = "dashed", size = 0.8) +
  geom_vline(aes(xintercept = mean_v - sd_v), color = "darkred", linetype = "dashed", size = 0.8) +
  #annotate("text", x = 1.5, y = 400, label = paste("mean:", round(mean_v, 2), 
  #                                                  "\nSD:", round(sd_v, 2)),
  #         hjust = 0.8, size = 2, color = "grey10") +
  labs(title = expression("tree vol."<= 2~m^3), x = expression(volume~(m^3)), y = "") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlim(0,2)

t4 <- ggplot(tree_mets, aes(v)) +
  geom_histogram(bins = 100) +
  labs(title = expression("tree vol."> 2~m^3), x = expression(volume~(m^3)), y = "") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlim(2, max(tree_mets$v))

cowplot::plot_grid(t1, t2, t3, t4, nrow = 1)

#ggsave("tree_mets_hists.jpg", plot = last_plot(), width = 7.5, height = 3, dpi = 500, device = "jpeg")
```


### by elevation 
```{r}
#tree heights
t1 <- ggplot(tree_mets, aes(x = altitude, y = h, color = habitat)) +
  geom_jitter(width = 5) +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) + 
  labs(title = "tree height", x = "elevation (m)", y = "height (m)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

#tree dbh
t2 <- ggplot(tree_mets, aes(x = altitude, y = dbh, color = habitat)) +
  geom_jitter(width = 5) +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) + 
  labs(title = "tree dbh", x = "elevation (m)", y = "diameter (cm)") + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none")

#tree volume
t3 <- ggplot(tree_mets, aes(x = altitude, y = v, color = habitat)) +
  geom_jitter(width = 5) +
  geom_smooth(se = TRUE, aes(color = NULL)) +
  scale_color_manual(values = ft_colors) + 
  labs(title = expression("tree vol."<= 2~m^3), x = "elevation (m)", y = expression(volume~(m^3))) + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none") + 
  ylim(0, 2)

#tree volume
t4 <- ggplot(tree_mets, aes(x = altitude, y = v, color = habitat)) +
  geom_jitter(width = 5) +
  scale_color_manual(values = ft_colors) + 
  labs(title = expression("tree vol."> 2~m^3), x = "elevation (m)", y = expression(volume~(m^3))) + 
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position = "none") + 
  ylim(2, max(tree_mets$v))

cowplot::plot_grid(t1, t2, t3, t4, nrow = 1)

#ggsave("tree_mets_elev.jpg", plot = last_plot(), width = 7.5, height = 3, dpi = 500, device = "jpeg")
```
